\ifdefined\anonymousbuild
\else
  \def\anonymousbuild{1}
\fi

\documentclass[conference]{IEEEtran}

\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{booktabs}
\usepackage{cite}
\usepackage{hyperref}
\usepackage{pgfplots}
\usepackage{xcolor}
\usepackage{listings}
\pgfplotsset{compat=1.18}

\newif\ifanonymous
\ifnum\anonymousbuild=1
  \anonymoustrue
\else
  \anonymousfalse
\fi

\input{generated/metrics.tex}

\newcommand{\toolname}{Ethereum Verification Benchmark}
\newcommand{\lean}{Lean 4}
\newcommand{\repoanon}{the public benchmark artifact}
\ifanonymous
  \newcommand{\artifactrepo}{\repoanon}
\else
  \newcommand{\artifactrepo}{\url{https://github.com/lfglabs-dev/ethereum-verification-benchmark}}
\fi

\lstdefinelanguage{LeanFour}{
  morekeywords={theorem,def,by,have,show,exact,simp,omega,import,namespace,end,structure,axiom,sorry,admit},
  sensitive=true,
  morecomment=[l]{--},
  morecomment=[n]{/-}{-/},
  morestring=[b]",
}
\lstset{
  language=LeanFour,
  basicstyle=\scriptsize\ttfamily,
  keywordstyle=\bfseries,
  breaklines=true,
  frame=single,
  columns=fullflexible,
  keepspaces=true
}

\makeatletter
\def\abstract{\normalfont
    \if@twocolumn
      \@IEEEabskeysecsize\bfseries\textit{\abstractname}:\relax
    \else
      \bgroup\par\addvspace{0.5\baselineskip}\centering\vspace{-1.78ex}\@IEEEabskeysecsize\textbf{\abstractname}\par\addvspace{0.5\baselineskip}\egroup\quotation\@IEEEabskeysecsize
    \fi\@IEEEgobbleleadPARNLSP}
\def\IEEEkeywords{\normalfont
    \if@twocolumn
      \@IEEEabskeysecsize\bfseries\textit{\IEEEkeywordsname}:\relax
    \else
      \bgroup\par\addvspace{0.5\baselineskip}\centering\@IEEEabskeysecsize\textbf{\IEEEkeywordsname}\par\addvspace{0.5\baselineskip}\egroup\quotation\@IEEEabskeysecsize
    \fi\@IEEEgobbleleadPARNLSP}
\makeatother

\title{Ethereum Verification Benchmark: An Open Suite for Evaluating AI Agents at Generating Lean 4 Proofs of Smart Contract Correctness}

\ifanonymous
\author{\IEEEauthorblockN{Anonymous Authors}}
\else
\author{\IEEEauthorblockN{Thomas Marchand}
\IEEEauthorblockA{Independent Researcher\\
Lisbon, Portugal\\
research@thomas.md}}
\fi

\begin{document}
\maketitle

\begin{abstract}
Formal verification is used in smart contract assurance, but mechanized proofs
remain expensive to write. Recent AI agents can use compilers and revise proof
scripts from error messages. This makes proof generation a concrete target
rather than a purely manual activity. Existing benchmarks rarely ask agents to
produce complete, machine-checkable proofs for contract obligations. They more
often test informal reasoning, small programs, or unit tests. \toolname{} is an
open evaluation suite for measuring whether AI agents can generate \lean{}
proofs of Ethereum smart contract correctness.
Built on Verity's shallow embedding of contract semantics in \lean{}, the
benchmark contains \BenchmarkTaskCount{} proof tasks across
\BenchmarkCaseCount{} deployed or ecosystem-relevant Ethereum cases. Each task
fixes the public proof interface and the verifier pipeline. Hidden reference
proofs are withheld from agents. The harness rejects \texttt{sorry},
\texttt{admit}, new axioms, theorem-statement edits, or unavailable imports.
Version manifests bind scores to fingerprints for the task and its execution
context. This makes reruns reproducible. We give the benchmark design and the
v0.1 release results. The strongest complete run, GPT-5.5, verifies 66 of 135
tasks. GLM 5.2 verifies 53 and Opus 4.8 verifies 45. Other complete runs range
down to 31 tasks. The strongest complete run solves less than half of the suite.
A budget-scaling sidecar shows that some failures depend on effort as well as
model choice.
\end{abstract}

\begin{IEEEkeywords}
smart contracts, formal verification, Lean 4, theorem proving, benchmarks,
AI agents, Ethereum
\end{IEEEkeywords}

\section{Introduction}

Formal verification of smart contracts has matured. Tools such as the Certora
Prover and K-based Ethereum analyses such as KEVM are used on production
protocols. Foundry has also made property tests part of everyday Ethereum
engineering practice~\cite{certora,kevm,kframework,foundry}.
The bottleneck is often not the kernel of the prover, but the human who
must translate protocol intent into precise invariants and then write the proof.
For large decentralized finance protocols, this labor can take weeks or
months and requires rare expertise at the intersection of protocol design,
EVM semantics, and theorem proving over fixed-width arithmetic.

AI agents make this bottleneck measurable in a new way. Modern agents can
use compiler feedback to revise proof scripts inside a workspace. In
mathematics and programming, benchmarks such as miniF2F, LeanDojo, HumanEval,
and SWE-bench have made progress easier to compare~\cite{minif2f,leandojo,humaneval,swebench}.
However, these benchmarks do not measure the target capability needed by
Ethereum verification teams: generating complete \lean{} proofs for
state-transition correctness properties of real smart contracts under a
machine-checked verifier.

\toolname{} targets this missing setting. It evaluates agents on proof-only
tasks derived from Ethereum protocols and ecosystem contracts. Each task gives
the agent a fixed contract model, a fixed formal specification, and one editable
\lean{} proof file. The agent succeeds only if an independent verifier accepts
the submitted proof. The verifier checks the theorem statement, rejects proof
placeholders and enforces a trusted axiom ledger. It then rebuilds the proof in
an isolated workspace. Reference proofs are not present in that workspace.

The benchmark is complementary to existing verification tools. Certora, KEVM,
K, Foundry, and similar tools help engineers state and check contract
properties. \toolname{} instead evaluates the proof-producing agent: can it
complete a formal proof once the contract semantics and property are fixed? It
is also distinct from general
software engineering and theorem-proving benchmarks because its tasks combine
uint256 arithmetic with protocol invariants drawn from deployed Ethereum
systems.

The paper contributes a benchmark protocol for AI-generated smart contract
proofs. The protocol isolates tasks, checks theorem integrity, rejects proof
placeholders, and enforces axiom accounting in CI. It also describes the v0.1
suite: \BenchmarkTaskCount{} tasks over \BenchmarkCaseCount{} cases and
\BenchmarkFamilyCount{} protocol families. Coverage is strongest in accounting,
solvency, storage effects, and linked-list ownership structures. The paper then
reports the v0.1 result artifacts. The best complete run solves 49\% of tasks.
The per-task outcomes show better performance on local arithmetic and storage
obligations than on longer invariant chains and cross-state reasoning.

\section{Related Work}

\textbf{Smart contract verification.}
The K framework provides executable semantics and program analysis
infrastructure, including KEVM for Ethereum bytecode~\cite{kframework,kevm}.
Runtime Verification applies these semantics to blockchain systems and
production audits~\cite{runtimeverification}. Certora provides a widely used
specification and verification workflow for smart contracts~\cite{certora}.
Foundry brings fast Solidity testing, fuzzing, and invariant tests into the
developer workflow~\cite{foundry}. These systems check contract behavior. Our
benchmark checks whether agents can synthesize the machine-checked proof
artifacts required by a \lean{}-based verification workflow.

\textbf{AI coding benchmarks.}
HumanEval measures function-level code synthesis from natural language
prompts~\cite{humaneval}. SWE-bench evaluates agents on issue-resolution tasks
in real software repositories~\cite{swebench}. These benchmarks measure coding
ability, but they do not require a formal proof object accepted by a theorem
prover, nor do they encode EVM-specific semantic obligations.

\textbf{LLMs and theorem proving.}
Benchmarks such as miniF2F and LeanDojo evaluate proof search in Lean and
related environments. Systems such as DeepSeek-Prover improve proof search in
this setting~\cite{minif2f,leandojo,deepseekprover}.
Their tasks are primarily mathematical. \toolname{} targets a different
distribution: contract state machines with storage-frame properties, arithmetic
rounding, authorization, solvency, and protocol-specific accounting.

\section{Background: Verity and Lean Proof Tasks}

Verity is a \lean{} framework for writing specifications and proofs for smart
contracts. It uses a shallow embedding of contract semantics. In a shallow
embedding, contract operations are represented using host-language definitions
and propositions rather than as uninterpreted syntax alone. This gives proof
authors direct access to \lean{}'s theorem prover while preserving a semantic
model of contract storage and state transitions.

A benchmark task fixes the implementation side and the theorem to prove. The
contract model defines the operation being verified. The specification states
the intended postcondition or invariant. The target theorem connects the
implementation to the specification.
The editable proof file initially contains the theorem statement and an empty
proof body; hidden reference proofs exist only for benchmark validation and are
not present in the agent workspace.

\begin{lstlisting}[caption={Schematic benchmark task shape.},label={lst:task}]
theorem deposit_sets_total_assets
  (s : VaultState) (assets : UInt256)
  (h : assets > 0) :
  spec_total_assets (deposit s assets h) =
    s.totalAssets + assets := by
  -- agent must complete this proof
\end{lstlisting}

The benchmark does not claim that every Ethereum behavior is fully modeled.
Instead, each case declares the formal surface it exercises, and the suite makes
coverage limitations explicit.

\section{Benchmark Design}

\subsection{Task Contract}

Every task is a proof-only problem. Agents may edit only designated proof files.
They may inspect the public task files. They also receive helper libraries and
a generated summary.
They may not see the hidden reference proof, modify the theorem statement, add
axioms, or alter the verifier. This keeps scores focused on proof synthesis
rather than benchmark adaptation.

\subsection{Axiom Ledger and Placeholder Rules}

The harness rejects \texttt{sorry}, \texttt{admit}, and new \texttt{axiom}
declarations in submitted proof files. Existing trusted assumptions are tracked
through a repository-level axiom ledger. CI enforces that the ledger remains
explicit and that proof files cannot silently expand the trusted computing base.
This is essential for a benchmark where otherwise an agent could ``solve'' a
task by assuming the target theorem.

\subsection{Theorem-Statement Integrity}

A proof attempt is not successful merely because Lean compiles. The verifier
checks that the theorem statement remains byte-for-byte consistent with the
task interface, modulo permitted proof-body edits. This prevents agents from
weakening preconditions, changing postconditions, or proving a related but
easier theorem.

\subsection{Isolated Workspaces}

Each run occurs in a fresh workspace containing public task files and generic
libraries, but not hidden proofs, private build artifacts, credentials, or
non-public imports. The final verifier then rebuilds the submitted proof in a
separate private copy. This two-step design reduces leakage from cached state
and catches imports that were not available to the agent.

\subsection{Verifier Pipeline}

The verifier checks the workspace policy and theorem integrity. It then checks
placeholder rejection, axiom compliance, import validity, and \lean{}
compilation. A task result is counted as verified only if all checks pass. Other
outcomes are recorded as failures or invalid artifacts. This keeps provider
failures and zero-token runs visible without ranking them as complete results.

\section{Suite Composition}

Version \BenchmarkVersion{} contains \BenchmarkTaskCount{} active task
manifests over \BenchmarkCaseCount{} cases and \BenchmarkFamilyCount{} protocol
families. The tasks are drawn from Ethereum protocols and ecosystem contracts,
including security challenge cases. The suite intentionally mixes simple local
storage updates with harder arithmetic and invariant-preservation tasks.

The active difficulty distribution is \BenchmarkEasyTasks{} easy,
\BenchmarkMediumTasks{} medium, \BenchmarkHardTasks{} hard, and
\BenchmarkOtherDifficultyTasks{} other. Stronger areas include accounting
conservation, local state preservation, guarded solvency, storage writes,
access-control identity, and linked-list ownership structures. Thinner areas in
v0.1 include deep reentrancy reasoning, oracle manipulation, governance,
timelock properties, liveness, cross-contract composition, cryptographic
assumptions, and adversarial EVM behavior.

\input{generated/suite_tables.tex}

\section{Methodology}

\subsection{Versioned Manifests}

Scores are comparable only relative to a fixed benchmark version. Version
\BenchmarkVersion{} records the ordered task set, Git commit, mode, budget,
and compatibility fingerprints. The task set identifier is
\BenchmarkTaskSetId{}. The harness identifier is \BenchmarkHarnessId{}. The
environment identifier is \BenchmarkEnvironmentId{}. The source commit recorded
by the manifest is \BenchmarkGitSha{}.

Each task also has a task fingerprint over execution-relevant files and fields,
and a task-interface fingerprint over the public files and theorem interface
visible to models. This separation allows future releases to distinguish
changes that require rerunning all tasks from changes that only affect a subset.

\subsection{Rerun Planning and Leaderboards}

The rerun planner compares old and new manifests. Harness, mode, budget, or
environment changes force broad reruns. Task additions, removals, or changed
fingerprints affect only the relevant task subset when compatible. Published
leaderboards separate complete rows from partial rows, so incomplete model runs
remain visible without being ranked against complete-suite runs.

\subsection{Metrics}

The primary metric is verified tasks over total active tasks. Token counts are
reported when providers expose them. They are secondary because provider
settings and agent wrappers differ. The v0.1 release separates complete-suite
runs from partial artifacts, so the
headline comparison ranks only complete runs while retaining partial runs as
diagnostic evidence.

\section{Preliminary Results}

Table~\ref{tab:results} reports every complete-suite v0.1 run published in the
release artifacts. GPT-5.5 verifies 66 of 135 tasks, the strongest complete
score in the release. GLM 5.2 verifies 53 tasks, Opus 4.8 verifies 45,
grok-build-0.1 verifies 41, MiniMax M3 verifies 38, and Kimi K2.7 verifies 31.
No complete run crosses the halfway point. The spread matters: current agents
do not fail uniformly. Still, the best result leaves 69 tasks unverified.
The suite therefore measures partial progress without treating proof generation
as solved.

\input{generated/result_figures.tex}
\input{generated/result_tables.tex}

Table~\ref{tab:partial-results} reports the non-complete artifacts from the
same v0.1 release. It includes partial and invalid runs. The most notable
partial artifact is GPT-5.5 Pro, which verifies 35 of 54 attempted tasks. That
is a high within-run rate, but it is not a complete-suite result and therefore
should not be ranked against complete 135-task runs. Invalid zero-usage
artifacts are retained in the release manifest because they document failed
provider or scheduling attempts rather than proof capability.

The MiniMax budget-scaling sidecar in Figure~\ref{fig:minimax-budget} answers a
different question from the leaderboard. It does not rerank models on the full
benchmark. Instead, it selects a 50-task panel consisting of the 38 tasks solved
by MiniMax M3 in the complete v0.1 run, topped up with 12 tasks solved by
GPT-5.5 but not MiniMax, and reruns only the still-unsolved tasks at increasing
budgets. On this selected panel, MiniMax reaches 48 of 50 tasks at the high
budget profile, or 96\%. This indicates that some apparent model gaps are
budget-sensitive. The experiment is reported as a sidecar because its adaptive
cascade methodology is not the same as the complete-suite leaderboard.

Manual inspection suggests that agents are strongest on local state updates,
direct arithmetic equalities, and proofs where the needed lemmas are close to
the target theorem. Failures often involve larger invariant chains. They also
involve non-obvious arithmetic normalization, stateful reasoning across several
fields, or finding the right helper lemma. This aligns with the suite
composition. Local accounting and storage obligations are better represented and
currently easier. Compositional protocol reasoning remains a future target.

\section{Discussion and Limitations}

\toolname{} measures proof synthesis under a fixed formalization. It does not
measure whether an agent can discover the right specification, identify an
unknown vulnerability, or decide whether the formal model matches the deployed
bytecode. Those are separate assurance problems. The benchmark is therefore best
understood as one layer of an AI-assisted verification stack: once engineers
choose the property and formal model, can an agent discharge the proof?

The v0.1 suite is also incomplete as a representation of Ethereum security.
It covers accounting, storage, access control, and solvency properties. It has
less coverage of reentrancy, oracle manipulation, governance, L2 messaging,
zero-knowledge protocol integration, cryptographic primitives, and
multi-contract composition. The immediate roadmap is to expand these families
while preserving the same verifier discipline and manifest-based comparability.

The model comparison should be read cautiously. The release artifacts record
per-task traces and hashes. They also record token totals and run caveats.
Providers expose different sampling controls and execution environments. For
that reason, v0.1 reports verified-task counts rather than a universal ranking
of models.

\section{Conclusion}

AI agents can already discharge some proof obligations in formal smart contract
verification, but the capability needs to be measured with complete,
reproducible tasks that reject shortcut solutions. \toolname{} provides such a
benchmark for \lean{} proofs of Ethereum contract correctness. Its v0.1 suite
contains 135 tasks from 25 real or ecosystem-relevant cases, enforces
no-placeholder proof rules, records compatibility fingerprints, and publishes
per-model artifacts. The v0.1 results give a concrete baseline: the best
complete run verifies 66 of 135 tasks. Several classes of protocol reasoning
remain mostly unsolved.

\ifanonymous
\else
\section*{Acknowledgment}
The public benchmark repository is available at \artifactrepo. The v0.1 release
artifacts were published on June 16, 2026.
\fi

\bibliographystyle{IEEEtran}
\bibliography{references}

\end{document}
